---
title: "Prac_Mach_Learn_2"
author: "N3P2"
date: "16 October 2015"
output: html_document
---

## Introduction  
Using health tracking devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.  

Data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants were used to predict the manner in which they did the exercise.  

## Data Preprocessing  

```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```

### Data download
```{r}
training.file   <- 'pml-training.csv'
test.cases.file <- 'pml-test.csv'
training.url    <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.cases.url  <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

download.file(training.url, training.file)
download.file(test.cases.url,test.cases.file)
```  

### Data read
Loading the data and reading the data sets on 2 frames

```{r}
trainRaw <- read.csv("pml-training.csv")
testRaw <- read.csv("pml-test.csv")
dim(trainRaw)
dim(testRaw)
```
The *training data set* contains **19622 observations and 160 variables**, while the *testing data set* contains **20 observations and 160 variables**. 

### Data cleaning
## Cleaning Data
All blank('""'), '#DIV/0' and 'NA' values are converted to 'NA'.
Columns containing 'NA' were removed from both data sets.
```{r}
training.df   <-read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
test.cases.df <-read.csv("pml-test.csv" , na.strings=c("NA", "#DIV/0!", ""))
training.df<-training.df[,colSums(is.na(training.df)) == 0]
test.cases.df <-test.cases.df[,colSums(is.na(test.cases.df)) == 0]
```

The first 7 columns are not related to calculations thus removed from data sets.
1. *user_name*
2. *raw_timestamp_part_1*
3. *raw_timestamp_part_2*
4. *cvtd_timestamp*
5. *new_window*
6. *num_window* 
```{r}
training.df   <-training.df[,-c(1:7)]
test.cases.df <-test.cases.df[,-c(1:7)]
```
Now, the cleaned training data set contains **19622 observations and 53 variables**, while the testing data set contains **20 observations and 53 variables**. The "classe" variable is still in the cleaned training set.

### Slice the data
Cleaned training set was split into a pure training data set (70%) and a validation data set (30%). Validation data set was used to conduct cross validation.  
```{r}
set.seed(22519) # For reproducibile purpose

inTrain <- createDataPartition(training.df$classe, p = 0.75, list = F)
trainData <- training.df[inTrain, ]
testData  <- training.df[-inTrain, ]
```

## Data Modeling
A predictive model was fitted for activity recognition using **Random Forest** algorithm due to automatic selection of important variables and robustness to correlated covariates & outliers in general.  **5-fold cross validation** will be used when applying the algorithm.
```{r}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
Using *confusion matrix* to test the performance of the model on the validation data set.  

```{r}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```

```{r}
accuracy <- postResample(predictRf, testData$classe)
accuracy

error.rate <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
error.rate
```
So, the *estimated accuracy* of the model is **99.42%** and the estimated *out-of-sample error* is **0.58%**.

## Test Data Set Prediction
```{r}
result <- predict(modelRf, test.cases.df[, -length(names(test.cases.df))])
result
```  

## Figures
1. Correlation Matrix Visualization  
```{r}
corrPlot <- cor(trainData[, -length(names(trainData))])
corrplot(corrPlot, method="color")
```
2. Decision Tree Visualization
```{r}
treeModel <- rpart(classe ~ ., data=trainData, method="class")
prp(treeModel) # fast plot
```

## Coursera code for submission
```{r}
answers <- result
pml_write_files <- function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i], file=filename, quote=FALSE,
                row.names=FALSE, col.names=FALSE)
  }
}
pml_write_files(answers)
```